Determining Causes of Death with ChatGPT: A Case Study of Deaths from Verbal Autopsy in Sierra Leone from 2019-2022

### 2024-01-02 manuscript/manuscript.tex
Update

\author[1,4]{\fnm{Rajeev} \sur{Kamadod}}\email{rajeevk@kentropy.com}

@@

@@

\affil[4]{\orgname{Kentropy Technologies Pvt. Ltd.}, \orgaddress{\street{2nd Main Rd}, \city{Bengaluru}, \postcode{560 034}, \state{Bangalore}, \country{India}}}

@@

Most low-income countries have reported to have either no data or have registered less than 50\% of deaths in their country, where only 8\% of these deaths have a Cause of Death (COD) recorded \cite{worldhealthorganizationSCOREHealthData2021}. To increase the number of records with a COD, a method known as Verbal Autopsy (VA) is used to determine likely causes for deaths outside of healthcare facilities. VA involves trained interviewers that speak to those familiar with the deceased to gather information based on standard questionnaires and open narratives, while medical professionals evaluate the information to assign probable CODs \cite{worldhealthorganizationVerbalAutopsyStandards2023}.

>>>>

\author[1,5]{\fnm{Rajeev} \sur{Kamadod}}\email{rajeevk@kentropy.com}

@@

\author[1,4]{\fnm{Patrick} \sur{Brown}}\email{}

@@

\affil[4]{\orgdiv{Department of Statistical Sciences}, \orgname{University of Toronto}, \orgaddress{\street{700 University Avenue}, \city{Toronto}, \postcode{M5G 1Z5}, \state{Ontario}, \country{Canada}}}

\affil[5]{\orgname{Kentropy Technologies Pvt. Ltd.}, \orgaddress{\street{2nd Main Rd}, \city{Bengaluru}, \postcode{560 034}, \state{Bangalore}, \country{India}}}

@@

Reliable counts of deaths and diagnosis of their causes are crucial for public health planning and policy making --- guiding interventions and resource allocation to reduce premature deaths \cite{lawnMillionNeonatalDeaths2010,lassiCommunityBasedIntervention2015,liuExcessMortalityPersons2017,ewigCommunityacquiredPneumoniaEmergency2011}. However, most low-income countries have reported to have either no data on deaths or have registered less than 50\% of deaths in their country, where an even fewer 8\% of these registered deaths have a Cause of Death (COD) recorded \cite{worldhealthorganizationSCOREHealthData2021}. To fill in this gap, an alternative method known as Verbal Autopsy (VA) is used to determine likely causes for deaths at scale and outside of traditional healthcare facilities \cite{desavignyIntegratingCommunitybasedVerbal2017,thomasVerbalAutopsyHealth2018,rampatigeSystematicReviewStatistics2014}. VA involves trained surveyors that interview those familiar with the deceased to gather information using standardized questionnaires and open narratives, while medical professionals evaluate the gathered information to assign probable CODs \cite{worldhealthorganizationVerbalAutopsyStandards2023}.
